---
title: "Dota2_Analysis_and_Results"
author: "Louis MS"
date: "22/03/2017"
output: html_document
---

write.csv(CombinedDF,  file = "CombinedDF.csv")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(CaTools)

# This RPubs document publishes to here: http://rpubs.com/LMS/Dota2_Analysis_and_Results
```

This document takes the previous work done in 'Dota2_Data_Wrangling.Rmd' which produced the data frame 'CombinedDF'. Lets take a quick look as a reminder of the variables collected.

```{r quick look at CombinedDF}
# Read in 'CombinedDF.csv' file
CombinedDF <- read.csv(file = "CombinedDF.csv")

# Glimpse at 'CombinedDF'
glimpse(CombinedDF)
```

## Gold Accumulation - Linear Regression Models
In this section we will address the question:

"How do I finish with a good amount of gold as a proxy to being 'successful' in the game?"

let's first look at how gold compares to match duration.
```{r Duration vs Gold XY plot}
# XY plot of duration against gold
ggplot(CombinedDF, aes(x = duration, y = gold, col = WL)) +
  geom_point(size = 0.2)
```

Here we see quite a broad range of gold values for different durations. Lets look at some heroes individually to see if we can clean up the analysis. Lets take a look at heroes' popularity and the percentage of matches they tend to win.

```{r Hero populatiry and win percentage}
# bar plot for most popular hero names 
Hero_Bar <- CombinedDF %>%
  group_by(Name, Class, Role_Count, All_Roles) %>%
  tally(sort = TRUE) %>%
  rename(HN_Count=n)

# Heroes win percentage (HWin_Pct)
HWin_Pct <- CombinedDF %>%
  group_by(Name, WL) %>%
  tally() %>%
  spread(key = WL, value = n) %>%
  mutate(Hero_Win_Pct = round(Win/(Win + Loss), digits = 2))

Hero_Bar <- left_join(Hero_Bar, HWin_Pct, by = "Name")

# ggplot of heroes ordered by popularity and color by their win percentage
ggplot(Hero_Bar, aes(x = reorder(Name, -HN_Count), y = HN_Count, width = 0.8, fill = Hero_Win_Pct)) +
  scale_fill_gradient2(low = "dark blue", high = "red", mid = "white", midpoint = 0.5) +
  geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
  xlab("Heroes ordered by popularity") +
  ylab("Heroes Used Count")

```

This plot show a significant level of popularity over other heroes for the top two most popular; Windranger and Shadow Fiend. It is also noted that Windranger loses more than she wins (blue shading indicates a <0.5 value for 'Hero_Win_Pct') and Shadow Fiend wins more than he loses (red shading indication >0.5 value for 'Hero_Win_Pct'), albeit these are both relatively close to the white 0.5 value for 'Hero_Win_Pct'. Lets also call for the hero with the highest 'Hero_Win_Pct', a variable made by 'win/(win+loss)'.

```{r Heroes with highest Hero_Win_Pct}
# ggplot of heroes ordered by win percentage and color by their popularity
ggplot(Hero_Bar, aes(x = reorder(Name, -Hero_Win_Pct), y = Hero_Win_Pct, width = 0.8, fill = HN_Count)) +
  scale_fill_gradient2(low = "dark blue", high = "red", mid = "white", midpoint = 5000) +
  geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
  xlab("Heroes ordered by win percentage") +
  ylab("Heroes win percentage")

```

Here we see Omniknight has the highest win percentage of `r max(Hero_Bar$Hero_Win_Pct)` . We will also examine this hero to get an indication of how this might vary the model.

### Windranger Linear Model
Lets collect all data on Windranger and look at the duration vs gold XY plot.

```{r Windranger duration vs gold XY plot}
# Selecting all Windranger data
Windranger <- CombinedDF[which(CombinedDF$Name=="Windranger"),]

# Windranger 'duration' vs 'gold' XY plot
ggplot(Windranger, aes(x = duration, y = gold, col = WL)) +
  geom_point(size = 0.2)
```

This plot shows Windranger's match duration vs gold, coloured by Win/Loss. The Windranger data frame contains 17597 observations. Let's look at gold and duration's linear relationship for Windranger.

```{r Windranger linear model 1 - gold predicted by duration}
# first linear model where 'duration' is used to predict 'gold'
WR_LM1 <- lm(gold ~ duration, data=Windranger)
summary(WR_LM1)

```

Model 'WR_LM1' shows a low and positive regression cooefficient with both the intercept and duration having very low P values, indicating that we can definitely reject the null hypothesis that there is no correlation between these variables. Gold and duration do correlate slightly positively, however the very low R-Squared value of 0.02678 tells us that the model is not very good. This tells us that generally over time you are likely to have more gold in longer matches, but there is quite a lot of variability on how much gold for a given duration. Lets see if we can improve this model by including additional variables. Lets add the variable 'gold_per_min' to the model as this is directly linked to the gold variable and duration. Let's also include the 'Win' variable to see how this affects the model.

```{r linear model 2 - Gold predicted by duration combined with gold_per_min}
# second linear model where 'duration' combined with 'gold_per_min' is used to predict 'gold'
WR_LM2 <- lm(gold ~ duration * gold_per_min + Win, data=Windranger)
summary(WR_LM2)

```

Model 'WR_LM2' displays a very small regression coefficient for 'duration*gold_per_min' of 0.003. However all variables have low P value indicating that there is a relationship with 'gold', as would be expected. The faster a hero collects gold and the longer they have to do this the more gold they are likely to finish with. Interestingly though we see that duration has changed from a positive correlation of about 0.5 to a negative correlation of less than -1. This seems to idicate now that with the addition of 'gold_per_min' and 'Win' the much longer matches may have an overall negative impact on the amount of gold a hero finishes. We see a significant improvement in the R-squared value to 0.3763 indicating that combining 'gold_per_min' and 'Win' with 'duration' has really improved the model. 

Let's see if we can improve on this model further by adding the 'gold_spent' variable as this will most definietly have a significant impact of gold in pocket at match end.

```{r linear model 3 - Gold predicted by duration, gold_per_min and gold_spent}
# third linear model where 'duration' combined with 'gold_per_min' plus 'gold_spent' is used to predict 'gold'
WR_LM3 <- lm(gold ~ duration * gold_per_min + Win + gold_spent, data=Windranger)
summary(WR_LM3)

```

Model 'WR_LM3'  shows a very large improvement in R-squared to 0.6534 so that adding 'gold_spent' has improved the model significantly. Interestly the P value for the intercept has increased quite a bit, but still remains at a low value of 0.0239. 

When a hero In Dota 2 dies the amount of gold is directly tied to their level. Lets add deaths and combine it with xp_per_min. Here we use xp_per_min because it is a continious variable that represents well the discrete variable 'level'.

```{r linear model - Gold predicted by duration, gold_per_min, gold_spent and deaths combined with xp_per_min}
# fourth linear model where 'duration' combined with 'gold_per_min', 'gold_spent' and 'deaths' combined with 'xp_per_min' is used to predict 'gold'
WR_LM4 <- lm(gold ~ duration * gold_per_min + Win + gold_spent + deaths * xp_per_min, data=Windranger)
summary(WR_LM4)

```

Model 'WR_LM4' shows further improvement in R-squared to 0.7221, which indicates a further increase in quality of the model.

We will include 'hero_damage' and 'tower_damage' into the model here since killing heroes and towers are significant sources of gold that players are encouraged to chase.


```{r linear model - Gold predicted by duration, gold_per_min, gold_spent, deaths combined with xp_per_min, hero_damage and tower_damage}
# fifth linear model where 'duration' combined with 'gold_per_min', 'gold_spent', 'deaths' combined with 'xp_per_min', 'hero_damage' and ' tower_damage' is used to predict 'gold'
WR_LM5 <- lm(gold ~ duration * gold_per_min + Win + gold_spent + deaths * xp_per_min + hero_damage + tower_damage, data=Windranger)
summary(WR_LM5)
```

Here we see model 'WR_LM5' a little improvement in R-squared to 0.7233. This model has the best R-squared value yet and is possibly the closest we will get to modelling the amount of gold a hero will likely finish a match with. We also note that all variables used in the model are very low apart from 'deaths' when modelled on its own. When included with 'xp_per_min' it remains significant.

Lets check this model with other heroes selected for this section.

### Linear modelling Shadow Fiend
```{r linear model - Shadow Fiend}
# Selecting all Shadow Fiend data
ShadowFiend <- CombinedDF[which(CombinedDF$Name=="Shadow Fiend"),]

# linear model of Shadow Fiend where 'duration' combined with 'gold_per_min', 'gold_spent', 'deaths' combined with 'xp_per_min', 'hero_damage' and ' tower_damage' is used to predict 'gold'
WR_LM5 <- lm(gold ~ duration * gold_per_min + gold_spent + deaths * xp_per_min + hero_damage + tower_damage, data=ShadowFiend)
summary(WR_LM5)
```


### Linear modelling Omniknight
```{r linear model - Omniknight}
# Selecting all Omniknight data
Omniknight <- CombinedDF[which(CombinedDF$Hero_Names=="Shadow Fiend"),]

# linear model of Omniknight where 'duration' combined with 'gold_per_min', 'gold_spent', 'deaths' combined with 'xp_per_min', 'hero_damage' and ' tower_damage' is used to predict 'gold'
WR_LM5 <- lm(gold ~ duration * gold_per_min + gold_spent + deaths * xp_per_min + hero_damage + tower_damage, data=Omniknight)
summary(WR_LM5)
```

The modelling of Shadow Fiend and Omniknight indicate that they have broadly similar responses to that of the Windranger dataset.


#### Gold Accumulation Conclusions
All models generated above showed that there was reason to believe that duration, gold_per_min, gold_spent, deaths, xp_per_min, hero_damage and tower_damage all correlate with the amount of gold a hero will finish with based upon the P-values for each variable. The best model 'WR_LM5' achieved an R-squared value of 0.6947 which indicates that variance from the model is fairly good. We can conclude that to finish the game with high amounts of gold players should try to maximise their rates of aqcuiring gold and xp, avoid deaths, takedown enemy players and towers and spend gold wisely at the shop.

## Best ways to play - Logistic Regression Models

Let's take a look at logistic regression modelling for Windranger to predict the wins and losses Windranger has and see if key variables can be identified.

```{r Windranger Logistic Regression Modelling}
# set the seed point
set.seed(1234)

# Split the Windranger data set into Training and Testing sets
table(Windranger$Win)
SplitWR <- sample.split(Windranger, SplitRatio = 0.5)

TrainWR <- subset(Windranger, SplitWR == TRUE)
TestWR <- subset(Windranger, SplitWR == FALSE)

# build logistic regression model for Windranger using duration, gold_spent, gold_per_min, xp_per_min, kills, deaths, assists, hero_damage and unit_order_total
Windranger.LogM1 <- glm(Win ~ duration + gold_spent + gold_per_min + xp_per_min + kills + deaths + assists + hero_damage + unit_order_total, data = TrainWR, family = binomial)
summary(Windranger.LogM1)

```





```{r}
# build logistic regression model for Windranger using gold_spent, gold_per_min, xp_per_min, kills, deaths, assists, hero_damage and unit_order_total
Windranger.LogM2 <- glm(Win ~ gold_spent + gold_per_min + xp_per_min + kills + deaths + assists + hero_damage + unit_order_total, data = TrainWR, family = binomial)
summary(Windranger.LogM2)
```


```{r}
# build logistic regression model for Windranger using duration, gold_spent, gold_per_min, xp_per_min, deaths, assists, hero_damage and unit_order_total
Windranger.LogM3 <- glm(Win ~ duration + gold_spent + gold_per_min + xp_per_min + deaths + assists + hero_damage + unit_order_total, data = TrainWR, family = binomial)
summary(Windranger.LogM3)
```


```{r}
# build logistic regression model for Windranger using duration, gold_spent * gold_per_min, xp_per_min * kills, deaths, assists, hero_damage and unit_order_total
Windranger.LogM4 <- glm(Win ~ duration + gold_spent * gold_per_min + xp_per_min * kills + deaths + assists + hero_damage + unit_order_total, data = TrainWR, family = binomial)
summary(Windranger.LogM4)

```


```{r}
ggplot(Windranger, aes(x = duration, colour = WL)) +
  geom_density()
```

```{r}
ggplot(ShadowFiend, aes(x = duration, colour = WL)) +
  geom_density()
```

```{r Windranger Logistic Regression Modelling}
# set the seed point
set.seed(1234)

# Split the Windranger data set into Training and Testing sets
SplitSF <- sample.split(ShadowFiend, SplitRatio = 0.5)

TrainSF <- subset(ShadowFiend, SplitWR == TRUE)
TestSF <- subset(ShadowFiend, SplitWR == FALSE)

# build logistic regression model for Windranger using duration, gold_spent, gold_per_min, xp_per_min, deaths, assists, hero_damage and unit_order_total
ShadowFiend.LogM <- glm(Win ~ duration+ gold_spent + gold_per_min + xp_per_min + deaths + assists + hero_damage + unit_order_total, data = TrainSF, family = binomial)
summary(ShadowFiend.LogM)

```