---
title: "Dota 2 Capstone Project - Detailed Analysis"
author: "Louis MS"
date: "07/03/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
# library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
# library(caTools)

# detach(package:plyr, unload = TRUE)

# look to library TidyVerse

# This RPubs document publishes to here: http://rpubs.com/LMS/xxx

```

## **In-Depth Detailed Analysis of Combined Dota 2 Dataset**
In the introduction and the initial exploratory data analysis two distinct questions come through:

A) Which group of heroes should I learn first?
B) What strategy should I persue with these heroes?

To answer Question A we will need to study ...


Lets look at the different classes and roles each hero takes on againsts the heroes popularity (how often a hero is selected by players) and how well they perform. Lets gather the mean for all hero stats when they win and when they lose. This will provide two rows per per hero.

### **Hero_Names Details and Figures **

```{r Hero_Bimod }
# collect averages for each hero split out by when they win and lose
Hero_Bimod <- CombinedDF %>% 
  select(Hero_Names, WL, level, kills, deaths, assists, denies, last_hits, hero_damage, gold, gold_per_min, xp_per_min, duration) %>% 
  rename(Name=Hero_Names) %>%
  mutate(KDdiff = Hero_KDR$kills - Hero_KDR$deaths) %>%
  group_by(Name, WL) %>%
  summarise(total_kills = sum(kills), 
            total_deaths = sum(deaths), 
            avg_kills = round(mean(kills), digits = 0), 
            avg_deaths = round(mean(deaths), digits = 0), 
            avg_levels = round(mean(level), digits = 0), 
            avg_assists = round(mean(assists), digits = 0), 
            avg_denies = round(mean(denies), digits = 0), 
            avg_las_hits = round(mean(last_hits), digits = 0),
            avg_hero_damage = round(mean(hero_damage), digits = 0), 
            avg_gold = round(mean(gold), digits = 0),
            avg_gold_per_min = round(mean(gold_per_min), digits = 0),
            avg_xp_per_min = round(mean(xp_per_min), digits = 0), 
            avg_KDdiff = round(mean(KDdiff), digits = 2),
            avg_duration = round(mean(duration), digits = 0) ) %>%
  mutate(KDR = round(total_kills/total_deaths, digits = 2))

# lets take a look at these averages when split out as win and loss
ggplot(data = melt(Hero_Bimod), aes(x = value, y = WL)) + 
  geom_point() + 
  facet_wrap(~variable, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```


Lets examine gold plotted against duration to see if a linear relationship can be defined. It is expected that as matches last longer, players should have more gold at the end of the match.

```{r gold v duration plot}
# xy plot of duration vs gold
ggplot(CombinedDF, aes(x = duration, y = gold, col = WL)) +
  geom_point(size = 0.2)

```

This plot shows the correlation between gold and duration is not clear. Lets look at the linear model of these two, and see if other variables can help the predictor.

```{r}
# examining linear model between duration and gold
model1 <- lm(gold ~ duration, data=CombinedDF)
summary(model1)
model1

# we can reject that there is no correlation (reject null hypothesis) but this is not a great model as indicated by the very low r^2 value. Lets see if the model improves

# here we see the model has very low P values which means we can reject the null hypothesis that these two variables, have no correlation, but the R^2 value of 0.02127 is very low indicating vey high variance.

# lets add WL as a predictor
model2 <- lm(gold ~ duration + WL, data=CombinedDF)
summary(model2)

# This improves R^2 significantly but it is still quite low at 0.2873. 

# Lets add Hero_Names as a predictor
model3 <- lm(gold ~ duration + WL + Hero_Names, data=CombinedDF)
summary(model3)

# lets add kills as a predictor but remove Hero_Names
model4 <- lm(gold ~ duration + WL + kills, data=CombinedDF)
summary(model4)

# lets add gold_death as a predictor
model5 <- lm(gold ~ duration + WL + kills + gold_death, data=CombinedDF)
summary(model5)

# r^2 is regression variavnce / total variance
```

```{r}
SSE1 <- sum(model1$residuals^2) # calculate sum of squared errors for model1
SSE1
```

```{r gold_death v duration plot}
# xy plot of duration vs gold_death
ggplot(CombinedDF, aes(x = duration, y = gold_death, col = WL)) +
  geom_point(size = 0.2)

GD_lm <- lm(gold_death ~ duration, data=CombinedDF)
summary(GD_lm)

```


```{r}

# interesting comparision plot to see how these variables spread out when distilled by WL
ggplot(data = melt(CombinedDF), aes(x = value, y = WL)) + 
  geom_point() + 
  facet_wrap(~variable, scales = "free_x") + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) 
```

CORRELATION MATRIX
```{r Correlation Matrix}


```


```{r level v duration plot}
# xy plot of gold vs duration
ggplot(CombinedDF, aes(x = duration, y = level, fill = WL)) +
  geom_point(size = 1)

```

```{r}
# examining linear model between level and duration
model1 <- lm(level ~ duration, data=CombinedDF)
summary(model1)

SSE1 <- sum(model1$residuals^2) # calculate sum of squared errors for model1
SSE1
```

```{r}
#            %>%
  mutate(KDR = round(total_kills/total_deaths, digits = 2)) 

rm(H_Bimod)
# %>%
#  select(Name, WL, KDR) %>%
#  spread(key=WL, value=KDR) %>%
#  rename(KDR_Loss=Loss, KDR_Win=Win) %>%
#  arrange(desc(KDR_Win), desc(KDR_Loss))

H <- Hero_KDR %>%
  group_by(Name, WL) %>%
  summarise(total_kills = sum(kills), total_deaths = sum(deaths)) %>%
  mutate(KDR = round(total_kills/total_deaths, digits = 2)) %>%
  select(Name, WL, KDR) %>%
  arrange(desc(KDR))

# ggplot hero KDRs
ggplot(H, aes(x = reorder(Name, -KDR), y = KDR, width = 0.5, fill = WL)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1, size = 5)) +
  xlab("Heroes") +
  ylab("Hero KDR")


```

#### *Hero_Bar*
Hero_Bar is a file containig key data with respect to each Hero Name designed to produce an array of bar charts to help explain the data. These variables are:

Hero_Names (rename to "Name"), Class, Role_Count, kills, deaths, assists, denies, last_hits, hero_damage, level and WL

```{r Hero_Bar }
# bar plot for top 40 most popular hero names 
Hero_Bar <- CombinedDF %>%
  group_by(Hero_Names, Class, Role_Count, All_Roles) %>%
  tally(sort = TRUE) %>%
  rename(Name=Hero_Names, HN_Count=n)

# Heroes win:loss ratio (WLR)

Hero_WLR <- CombinedDF %>%
  group_by(Hero_Names, WL) %>%
  tally() %>%
  spread(key = WL, value = n) %>%
  rename(Name=Hero_Names) %>%
  mutate(WLR = round(Win/Loss, digits = 2))

Hero_Bar <- left_join(Hero_Bar, Hero_WLR, by = "Name")
```

```{r Bar plot of Hero Name Count colour filled by WLR}
# ggplot to show how hero selection stands up to their match Win:Loss ratio
ggplot(head(Hero_Bar, n = 40), aes(x = reorder(Name, -HN_Count), y = HN_Count, width = 0.8, fill = WLR)) +
  scale_fill_gradient2(low = "dark blue", high = "red", mid = "white", midpoint = 1) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Role_Count, vjust = 2)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
  xlab("Top 40 Most Popular Heroes") +
  ylab("Heroes Used Count")

```
This bar graph shows the popularity of the top 40 heroes by selection coloured by their win:loss ratio (WLR), where white represents 1:1 win:loss, red represents better than 1:1 win:loss and blue represents worse than 1:1 win:loss. The integer numbers in each bar represent the number of roles each hero is associated with, where higher numbers indicate a hero who is more versitile and able to carry out a wider range of functions. Please note the total number of all heroes counted is `r sum(Hero_Bar$HN_Count)`.

Let's see how these stack up when ordered by HN_Count as a percent and coloured by WLR.

```{r}
Hero_Bar$HN_Pct <- round((Hero_Bar$HN_Count/sum(Hero_Bar$HN_Count)*100), digits = 2)

ggplot(head(Hero_Bar, n = 40), aes(x = reorder(Name, -HN_Count), y = HN_Pct, width = 0.8, fill = WLR)) +
  scale_fill_gradient2(low = "dark blue", high = "red", mid = "white", midpoint = 1) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
  xlab("Top 40 Heroes") +
  ylab("Win:Loss Ratio")

```

Let's take a look at each heroes kill:death ratios (KDR) when they win and when they lose matches.

```{r}
# heroes kill:death ratio (KDR)
Hero_KDR <- CombinedDF %>% 
  select(Hero_Names, account_id, match_id, WL, level, kills, deaths) %>% 
  rename(Name=Hero_Names, Player=account_id) %>%
  mutate(KDR = round(Hero_KDR$kills / Hero_KDR$deaths, digits = 2)) %>%
  mutate(KDdiff = Hero_KDR$kills - Hero_KDR$deaths)

# summarise WL kills and deaths
HKDR <- Hero_KDR %>%
  group_by(Name, WL) %>%
  summarise(total_kills = sum(kills), total_deaths = sum(deaths)) %>%
  mutate(KDR = round(total_kills/total_deaths, digits = 2)) %>%
  select(Name, WL, KDR) %>%
  spread(key=WL, value=KDR) %>%
  rename(KDR_Loss=Loss, KDR_Win=Win) %>%
  arrange(desc(KDR_Win), desc(KDR_Loss))

H <- Hero_KDR %>%
  group_by(Name, WL) %>%
  summarise(total_kills = sum(kills), total_deaths = sum(deaths)) %>%
  mutate(KDR = round(total_kills/total_deaths, digits = 2)) %>%
  select(Name, WL, KDR) %>%
  arrange(desc(KDR))

# ggplot hero KDRs
ggplot(H, aes(x = reorder(Name, -KDR), y = KDR, width = 0.5, fill = WL)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 70, hjust = 1, size = 5)) +
  xlab("Heroes") +
  ylab("Hero KDR")
```

```{r}
# Look at WLR vs KDR in an XY plot
Hero_Bar <- left_join(Hero_Bar, HKDR, by = "Name")

ggplot(Hero_Bar, aes(x = KDR_Win, y = KDR_Loss, col = WLR)) +
  scale_color_gradient2(low = "blue", high = "red", mid = "white", midpoint = 1) +
  geom_point(size = 2)

```


```{r}
# look at gather
# p <- spread(HLWR, key = WL, value = (win_total_kills = total_kills")))


# rm(p)
gather(data = HLWR, key = Name, value = )


# group by player and tally
GP_Player <- Hero_KDR %>%
  group_by(Player) %>%
  tally(Player != 0, sort = TRUE)

max(GP_Player$n)
mean(GP_Player$n)

max(GP_Player$Player)

GPP_10 <- filter(GP_Player)

# ggplot of player count
# ggplot(head(GP_Player, n = 40), aes(x = Player, y = n)) + geom_bar(stat = "identity")


HKDR <- Hero_KDR %>% select(Name, WL, KDR)

kd <- aggregate(HKDR, FUN = mean, by = Name)


rm(kd)

# kd <- cast(HKDR, Name~KDR, mean)
# aggregate(data=HKDR, by=Name, FUN=mean)

# looking at Players


```


```{r}
# percent of tatal matches sampled
Hero_Bar$Pct <- round(Hero_Bar$HN_Count/sum(Hero_Bar$HN_Count)*100, digits=1)

# plot showing count and percent for each hero selected
ggplot(head(Hero_Bar, n=40), aes(x=reorder(Name, -HN_Count), y=HN_Count, width=0.8, fill=Class, as.factor=TRUE)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label=Pct, vjust = 0.5, hjust = 0), check_overlap = TRUE) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1, size = 8)) +
  xlab("Top 40 Most Popular Heroes") +
  ylab("Count")
```


```{r}
# by percent instead of count, with Role_Count displaying on bars
ggplot(head(Hero_Bar, n = 40), aes(x = reorder(Name, -HN_Count), y = Pct, width = 0.8, fill = Class, as.factor = TRUE)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Role_Count, vjust = 2)) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1, size = 8)) +
  xlab("Top 40 Most Popular Heroes") +
  ylab("Percent")

```

```{r Hero Name use Pct bar chart colored by WLR}

# ggplot of Hero_Bar Count coloured by WLR

ggplot(head(Hero_Bar, n = 40), aes(x = reorder(Name, -HN_Count), y = HN_Count, width = 0.8, fill = WLR)) +
  scale_fill_gradient2(low = "dark blue", high = "red", mid = "white", midpoint = 1) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Role_Count, vjust = 2)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
  xlab("Top 40 Most Popular Heroes") +
  ylab("Heroes Used Count")
```

This bar chart shows how many times each hero was selected coloured by their respective Win:Loss Ratio (WLR). The number within each bar represents the number of roles each hero is classified with.  For example Windranger (5) has been classified for Carry, Disabler, Support, Nuker and Escape roles, where as Shadow Fiend (2) has been classified for only Carry and Nuker roles. It is also noted that Shadow Fiend has an above par WLR 1.09 where as Windranger has a below par WLR 0.92 which opens for some early thories here. Windranger although more popular and diverse in roles has lost more matches than won, unlike the second most popular Shadow Fiend who only has two roles which are to get strong and deal massive burst damage. 

Here we see two interesting heroes immediately, Windranger and Shadow Fiend who were selected 4.2% and 3.4% of the time respectively. Beyond these two all other heroes in the top 40 were selected between 2.3% and 0.9% of the time.

```{r Looking at Windrunner and Shadow Fiend}
# print(Hero_Bar[Name = c("Windranger", "Shadow Fiend")])



```

```{r}
# bar plot for all hero names
ggplot(Hero_Bar, aes(x = reorder(Name, -HN_Count), y = HN_Count, width = 0.75, fill = Class, as.factor = TRUE)) +
  geom_bar(stat = "identity", aes()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, size = 5))
```


```{r Investigating CLass}
# total Class tally and plot of top 40 most used Heroes
Class_Ct <- Hero_Bar %>%
  group_by(Class) %>%
  tally(sort = TRUE)

# percent of tatal matches sampled
Class_Ct$Pct <- round(Class_Ct$n/sum(Class_Ct$n)*100, digits = 1)

ggplot(Class_Ct, aes(x = Class, y = Pct, width = 0.5, fill = Class, as.factor = TRUE)) +
  geom_bar(stat = "identity")
Class_Ct


```

```{r}

Cols <- c("1.45" = "red", "1.01" = "red", "1" = "black", "0.99" = "steel blue", "0.68" = "steel blue")

ggplot(Hero_WLR,aes(x = reorder(Name, -WLR), y = Win, fill = WLR, as.factor = TRUE)) +
  scale_fill_manual(values = Cols, breaks = c("1.45", "1.01", "1", "0.99", "0.68")) +
         geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, size = 5))



  scale_fill_gradient2(low = "steel blue", high = "red", mid = "white", midpoint = 1) +

min(Hero_WLR$WLR)
```





```
