---
title: "Testing_Linear_Models"
author: "Louis MS"
date: "27/03/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r} 
ggplot(Windranger, aes(x = xp_per_min, y = gold, col = WL)) +
  geom_point(size = 0.2)
```


Variables to test against gold:
duration      (9)
gold_spent    (7)
gold_per_min  (1)
xp_per_min    (3)
kills         (4)
deaths        (8)
level         (5)
hero_damage   (6) 
tower_damage  (2)

Number refers to the highest R-squared value of each model when modelly gold vs single variable.

```{r gold vs ...}
LM1 <- lm(gold ~ duration * gold_per_min + tower_damage + hero_damage + deaths * xp_per_min + gold_spent, data=Windranger)
summary(LM1)

```

R-squared = 0.02678

```{r gold v gold_spent}
LM2 <- lm(gold ~ gold_spent, data=Windranger)
summary(LM2)
```

R-squared = 0.07571

```{r gold v gold_per_min}
LM3 <- lm(gold ~ gold_per_min, data=Windranger)
summary(LM3)
```

R-squared = 0.2604

```{r gold v xp_per_min}
LM4 <- lm(gold ~ xp_per_min, data=Windranger)
summary(LM4)
```

R-squared = 0.1797

```{r gold v kills}
LM5 <- lm(gold ~ kills, data=Windranger)
summary(LM5)
```

R-squared = 0.1434

```{r gold v deaths}
LM6 <- lm(gold ~ deaths, data=Windranger)
summary(LM6)
```

R-squared = 0.07458

```{r gold v level}
LM7 <- lm(gold ~ level, data=Windranger)
summary(LM7)
```

R-squared = 0.1382

```{r gold v hero_damage}
LM8 <- lm(gold ~ hero_damage, data=Windranger)
summary(LM8)
```

R-squared = 0.1

```{r gold v tower_damage}
LM9 <- lm(gold ~ gold_spent, data=Windranger)
summary(LM9)
```

R-squared = 0.2502

Windranger.lm <- lm(gold ~ duration * gold_per_min + Win + gold_spent + deaths * xp_per_min + hero_damage, data=Windranger)
summary(WR_LM5)

